---
title: "a3_task3_hollie_pennington"
author: "Hollie Pennington"
date: "10/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(kableExtra)
library(lubridate)
library(janitor)
library(plotly)
```

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names
```

#Section 1: Size Comparisons: purple and red urchins 

```{r}
ggplot(data = urchins, aes(x = size)) + #read in data for ggplot
  geom_histogram(bins = 10) + #choose graph sytle and bin number
  facet_wrap(~ common_name) + #faceted by color (common name)
  labs(x = "Urchin Diameter (cm)", title = "Purple VS Red Urchins Size Comparison", caption = "Figure 1: Shows the distribution of size in diameter (cm) in purple sea urchins compared to \n red sea urchins.") + #labeled graphs
  theme_minimal()
```
The purple urchin data is skewed to the right. The red urchin data has a semi normal distribution. 

#Descriptive Statistics Table 

```{r}
urchins_table <- urchins %>% #name table from urchins data
  group_by(common_name) %>%  #group by the common name 
  summarise(mean_size = mean(size, na.rm = TRUE), #find mean 
            sd_size = sd(size, na.rm = TRUE), #find standard deviation
            sample_size = n()) #find sample size 

urchins_table #show table 

urchins_table %>% #let kableExtra know where to pull data from 
  kable(col.names = c("Species", "Mean Diameter (cm)", "Standard Deviation (cm)", "Samlple Size (n)"), caption = "<b> Table 1 </b> Table compares mean diameter in cm, standard deviation in cm, and sample size number between purple and red urchins") %>% #named columns and added caption
  kable_styling(full_width = FALSE) #fixed width 
```
The purple urchin's data has a mean diameter is around 4.47 cm with a standard deviation of 1.2 cm from a population of 31805. The red urchin's data has a mean diameter of 6.78 cm with a standard deviation of 1.9 cm from a population size of 27408. The red urchins have a larger mean diameter than the purple urchins. 

#Section 2

```{r}
purple_urchins <- urchins %>% #created new_named data set
  filter(common_name == "Purple Urchin") %>% #filtered for purple urchins
  mutate(date = mdy(date)) %>% #created date 
  mutate(year = year(date)) %>% # create new column with year 
  mutate(Site = case_when(
    site %in% c("NAPL") ~ "Naples Reef", 
    site %in% c("AQUE") ~ "Arroyo Quemado Reef",
    site %in% c("CARP") ~ "Carpinteria Reef",
    site %in% c("IVEE") ~ "Isla Vista Reef"
  )) #change site names from abbreviations to full names
```

```{r}
purple_urchins_table <- purple_urchins %>% #create new data frame
  group_by(Site, year) %>% #grouped for site and year
  summarize(total = n()) #found total for year and site 
```

```{r}
ggplot(data = purple_urchins_table, aes(x = year, y = total, fill = Site)) +
  geom_col(position = position_dodge(), width = 0.75) + 
  labs(title = "Total Purple Urchins", x = "Year", y = "Number of Urchins", caption = "Figure 2: Graph showing total number of purple urchins found in various sites from years \n 2008 to 2020. Sites include Arroyo Quemado Reef, Caprinteria Reef, Isla Vista Reef, \n Naples Reef, and not recoreded on a beach. The year 2009 had the highest number \n of total urchins in every beach site. Carpinteria Reef was the single beach with the \n higest recorded purple urchin count.") + 
  theme(legend.position = c(0.8, 0.7), 
        legend.key.size = unit(.3, 'cm'), )
```
The data in this graph shows a semi normal distribution with a slight skew to the right. The years 2009, 2010, and 2011 had the highest counts of purple urchins. 